\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-04-24 doc/paper.tex
added paragraph at end of introduction describing paper structure. Changed this structure a little.

Detailed phenonenological studies have been carried out for scenarios
within the minimal supersymmetric standard model (MSSM).  Such work
has been greatly aided by public spectrum generators for
MSSM\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae},
allowing fast and reliable exploration of the sparticle spectrum and
mixings that can be obtained from particular choices of breaking
mechanism inspired boundary conditions and specified parameters. There
are also two public spectrum generators
\cite{Ellwanger:2006rn,Allanach:2013kza} for the next to minimal
supersymmetric standard model (NMSSM) \cite{NMSSM} .

None of the fundamental motivations of supersymmetry require
minimality, and specific altternatives to (or extensions of) the MSSM
are motiovated by, for example,the $\mu-problem$ of the
MSSM\cite{Kim:1983dt}, explaining the family structure (see e.g,
\cite{King:2014nza}) or for successful baryogensis or leptogenesis
(see e.g,\cite{King:2008qb}). However constructing specialised tools
to study all relevant models would require an enormous amount of work.
So general tools which can automate this process and produce fast and
reliable programs can greatly enhance our ability to understand and
test non-minimal realisations of supersymmetry.

@@

These motivate both exploring non-minimal SUSY models which ameliorate
the naturaless problems introduced, by, for example, raising the tree
level Higgs mass, and models developed with a fresh persp[ective,
  based on other considerations.  In both cases exploration of such
  models can be aided if it is possible to quickly create a fast
  spectrum generator.

\fs is designed for precisely this task. \fs is a MATHEMATICA package
which with the help of \sarah
\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}
creates a fast, modular C++ spectrum generator for a user specified
SUSY model.  The generated code structure is designed to be as
flexible as possible to accomodate different types of extensions and
due to it's modular nature it is easy to modify, add new featuires to
or combine with other programs.

\subsection{The Program}
\label{Sec:Program}
\fs is a MATHEMATICA package designed to create a fast and easily
adaptable spectrum generator in C++ for any SUSY model and some
non-supersymmetric models {\color{red} can we?}.  \fs makes use of the
existing \sarah package \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} to obtain one-loop self energies, one-loop
tadpoles corrections and two-loop renormalisation group
equations(RGEs) in any model.  \fs then converts these algerbraic
expressions from \sarah into C++ code using the C++11 standard. \fs
uses some parts of SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen
library \cite{eigen}, as well as the gnu scientific library and the
boost library to create numerical routines which then use these
routines to solve the RGEs and EWSB conditions which will be used in
the spectrum generator. The main solver routine of \fs then calls
these routines to find simultaneous solutions for the EWSB, low energy
data, and other, user supplied, boundary conditions (e.g. the usual
mSUGRA universality conditions) and calculates the mass spectrum.

@@

\section{Download and compilation}
Flexible SUSY can be downloaded from \url{http://flexiblesusy.hepforge.org}.  
To use FlexibleSUSY please make sure you have the following installed.  
\begin{itemize}
\item Mathematica, version 7 or higher
\item SARAH, version 4.0.4 or higher \url{http://sarah.hepforge.org}
\item C++11 compatible compiler (g++ 4.4.7 or higher, clang++ 3.1 or
  higher)
\item Fortran compiler (gfortran, ifort etc.)
\item Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}
\item Boost library, version 1.36.0 or higher
  \url{http://www.boost.org}
\item GNU scientific library \url{http://www.gnu.org/software/gsl/}
\item lapack (needed for the Lattice algorithm only)
  \url{http://www.netlib.org/lapack/}
\end{itemize}
%
Optional:
%
\begin{itemize}
\item Looptools, version 2.8 or higher
  \url{http://www.feynarts.de/looptools/}
\end{itemize}

@@

\section{Setting up a FlexibleSUSY model}

A \fs model is configured via a (Mathematica) model file, which in
case of the MSSM looks like this
%
\begin{lstlisting}[language=Mathematica]
FSModelName = "MSSM";

@@

\section{Stucture of the spectrum generator}

\subsection{Model parameters and RGEs}
Even in the simplest supersymmetric model the RGEs form a very large
set of coupled non-linear differential equations.  In a general
(softly broken) supersymmetric theory the soft-breaking parameters
depend on themselves and the susy parameters.  However, the susy
parameters are independent of the soft-breaking parameters
\cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Sperling:2013eva,Sperling:2013xqa}.
This makes it possible to split the model parameters into the
following two subsets:
%
\begin{enumerate}
\item \emph{susy parameters:} gauge couplings, superpotential
  parameters and VEVs and
\item \emph{soft-breaking parameters} \cite{Girardello:1981wz}: soft
  linear scalar terms, soft bilinear scalar interactions, soft
  trilinear scalar interactions, soft gaugino mass terms and soft
  scalar squared masses.
\end{enumerate}
%
The susy parameters together with their two-loop renormalisation group
equations are stored in the class \code{<model>_susy_parameters}.  The
soft-breaking model parameters together with their two-loop
renormalisation group equations are stored in the class
\code{<model>_soft_parameters}.  The dependence of the soft-breaking
parameters on the susy parameters is implemented via inheritance, see
\figurename~\ref{fig:parameter-classes}, where
\code{<model>_soft_parameters} inherits from
\code{<model>_susy_parameters}.  In this way, the class
\code{<model>_soft_parameters} can implement the soft-breaking beta
functions in terms of the susy and soft-breaking parameters, while the
class \code{<model>_susy_parameters} is independent of the
soft-breaking parameters.
%
\begin{figure}
  \centering
  \tikzumlset{fill class=white}
  \begin{tikzpicture}
    \umlclass[x=0, y=3]{model\_susy\_parameters}{
      - susy parameters
    }{
      + beta()
    }
    \umlclass[x=0, y=0]{model\_soft\_parameters}{
      - soft-breaking parameters
    }{
      + beta()
    }
    \umlinherit{model\_soft\_parameters}{model\_susy\_parameters}
  \end{tikzpicture}
  \caption{Model parameter class structure.}
  \label{fig:parameter-classes}
\end{figure}

@@

\section{Implementing new models}
\section{Adapting the code at the C++ level}
\section{Using individual modules in your own code}

@@

>>>>

Detailed phenomenological studies have been carried out for scenarios
within the minimal supersymmetric standard model (MSSM).  Such work
has been greatly aided by public spectrum generators for
MSSM\cite{Allanach:2001kg,Porod:2003um,Djouadi:2002ze,Baer:1993ae},
allowing fast and reliable exploration of the sparticle spectrum and
mixings that can be obtained from particular choices of breaking
mechanism inspired boundary conditions and specified parameters. There
are also two public spectrum generators
\cite{Ellwanger:2006rn,Allanach:2013kza} for the next to minimal
supersymmetric standard model (NMSSM) \cite{NMSSM} .

None of the fundamental motivations of supersymmetry require
minimality, and specific alternatives to (or extensions of) the MSSM
are motivated by, for example,the $\mu-problem$ of the
MSSM\cite{Kim:1983dt}, explaining the family structure (see e.g,
\cite{King:2014nza}) or for successful baryogensis or leptogenesis
(see e.g,\cite{King:2008qb}). However constructing specialised tools
to study all relevant models would require an enormous amount of work.
So general tools which can automate this process and produce fast and
reliable programs can greatly enhance our ability to understand and
test non-minimal realisations of supersymmetry.

@@

These motivate both exploring non-minimal SUSY models which ameliorate
the naturalness problems introduced, by, for example, raising the tree
level Higgs mass, and models developed with a fresh perspective,
  based on other considerations.  In both cases exploration of such
  models can be aided if it is possible to quickly create a fast
  spectrum generator.

\fs is designed for precisely this task. \fs is a MATHEMATICA package
which with the help of \sarah
\cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta}
creates a fast, modular C++ spectrum generator for a user specified
SUSY model.  The generated code structure is designed to be as
flexible as possible to accommodate different types of extensions and
due to it's modular nature it is easy to modify, add new features to
or combine with other programs.  The generated code has been
extensively tested against well known spectrum generators and as well
providing a solution for new SUSY models the generated MSSM and NMSSM
codes offer a modern and fast alternative to the existing public
spectrum generators.

In section \ref{Sec:Program} we describe the program in more detail
and explain our design goals.  In section \ref{Sec:download}
information on how to download and compile the code may be found along
with details on how to get started quickly.  In section
\ref{Sec:modfile} we describe how the user can create a new
FlexibleSUSY model file. A detailed description of the structure and
features of the generated code is then given in section
\ref{Sec:SpecGenStruct}.  In section \ref{Sec:Flexible} we describe
the various ways the code can be modified both at the meta code level
by writing model files and at the C++ code level by modifying the code
or adding new modules. Finally in section \ref{Sec:comparison} we
describe detailed comparisons between our generated code and existing
public spectrum generators as well as against the SPheno-like fortran
code which can be created using SARAH.

\section{The Program}
\label{Sec:Program}
\fs is a MATHEMATICA package designed to create a fast and easily
adaptable spectrum generator in C++ for any SUSY model and some
non-supersymmetric models {\color{red} can we?}.  \fs makes use of the
existing \sarah package \cite{Staub:2010ty,Staub:2009bi,Staub:2010jh,Staub:2012pb,Staub:2013tta} to obtain one-loop self energies, one-loop
tadpoles corrections and two-loop renormalisation group
equations(RGEs) in any model.  \fs then converts these algerbraic
expressions from \sarah into C++ code using the C++11 standard. \fs
uses some parts of SOFTSUSY\cite{Allanach:2001kg}, the very fast Eigen
library \cite{eigen}, as well as the gnu scientific library and the
boost library to create numerical routines which then use these
routines to solve the RGEs and EWSB conditions which will be used in
the spectrum generator. The main solver routine of \fs then calls
these routines to find simultaneous solutions for the EWSB, low energy
data, and other, user supplied, boundary conditions (e.g. the usual
mSUGRA universality conditions) and calculates the mass spectrum.

@@

\section{Download and compilation}
\label{Sec:download}
Flexible SUSY can be downloaded from \url{http://flexiblesusy.hepforge.org}.  
To use FlexibleSUSY please make sure you have the following installed.  
\begin{itemize}
\item Mathematica, version 7 or higher
\item SARAH, version 4.0.4 or higher \url{http://sarah.hepforge.org}
\item C++11 compatible compiler (g++ 4.4.7 or higher, clang++ 3.1 or
  higher)
\item Fortran compiler (gfortran, ifort etc.)
\item Eigen library, version 3.1 \url{http://eigen.tuxfamily.org}
\item Boost library, version 1.36.0 or higher
  \url{http://www.boost.org}
\item GNU scientific library \url{http://www.gnu.org/software/gsl/}
\item lapack (needed for the Lattice algorithm only)
  \url{http://www.netlib.org/lapack/}
\end{itemize}
%
Optional:
%
\begin{itemize}
\item Looptools, version 2.8 or higher
  \url{http://www.feynarts.de/looptools/}
\end{itemize}

@@

\section{Setting up a FlexibleSUSY model}
\label{Sec:modfile}
A \fs model is configured via a (Mathematica) model file, which in
case of the MSSM looks like this
%
\begin{lstlisting}[language=Mathematica]
FSModelName = "MSSM";

@@

\section{Stucture of the spectrum generator}
\label{Sec:SpecGenStruct}
\subsection{Model parameters and RGEs}
Even in the simplest supersymmetric model the RGEs form a very large
set of coupled non-linear differential equations.  In a general
(softly broken) supersymmetric theory the soft-breaking parameters
depend on themselves and the susy parameters.  However, the susy
parameters are independent of the soft-breaking parameters
\cite{Jones:1974pg,Jones:1983vk,West:1984dg,Martin:1993yx,Yamada:1993ga,MV94,Fonseca:2011vn,Sperling:2013eva,Sperling:2013xqa}.
This makes it possible to split the model parameters into the
following two subsets:
%
\begin{enumerate}
\item \emph{susy parameters:} gauge couplings, superpotential
  parameters and VEVs and
\item \emph{soft-breaking parameters} \cite{Girardello:1981wz}: soft
  linear scalar terms, soft bilinear scalar interactions, soft
  trilinear scalar interactions, soft gaugino mass terms and soft
  scalar squared masses.
\end{enumerate}
%
The susy parameters together with their two-loop renormalisation group
equations are stored in the class \code{<model>_susy_parameters}.  The
soft-breaking model parameters together with their two-loop
renormalisation group equations are stored in the class
\code{<model>_soft_parameters}.  The dependence of the soft-breaking
parameters on the susy parameters is implemented via inheritance, see
\figurename~\ref{fig:parameter-classes}, where
\code{<model>_soft_parameters} inherits from
\code{<model>_susy_parameters}.  In this way, the class
\code{<model>_soft_parameters} can implement the soft-breaking beta
functions in terms of the susy and soft-breaking parameters, while the
class \code{<model>_susy_parameters} is independent of the
soft-breaking parameters.
%
\begin{figure}
  \centering
  \tikzumlset{fill class=white}
  \begin{tikzpicture}
    \umlclass[x=0, y=3]{model\_susy\_parameters}{
      - susy parameters
    }{
      + beta()
    }
    \umlclass[x=0, y=0]{model\_soft\_parameters}{
      - soft-breaking parameters
    }{
      + beta()
    }
    \umlinherit{model\_soft\_parameters}{model\_susy\_parameters}
  \end{tikzpicture}
  \caption{Model parameter class structure.}
  \label{fig:parameter-classes}
\end{figure}

@@

%%\section{Implementing new models}
\section{Flexible Spectra}
\label{Sec:Flexible}
\subsection{Meta code}
\subsection{Adapting C++ code}
\subsection{Combing with  modules of your own code}

@@

\label{Sec:comaprison}
